---
title: 'Task 2: Model selection'
author: 'By: Claire Gonzales'
date: "1/24/2022"
output: 
  html_document: 
    theme: lumen
---

### Project Overview

This project will use seawater sample data from CalCOFI research cruises to explore physical relationships. Ultimately, this report aims to discern the relationship between oxygen saturation levels and other physical and chemical variables. I will use AIC and cross validation to compare the preformance of two different linear regression models to find the model of best fit. 

(Data: CalCOFI 2021)


```{r setup, include= TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

# attach packages

library(tidyverse)
library(here)
library(lubridate)
library(AICcmodavg)
library(equatiomatic)
```

```{r}
# reading in csv data

seawater <- read_csv(here("data", "calcofi_seawater_samples.csv")) %>% 
  drop_na()
```

### Create 2 models

The two multiple linear regressiong models will address the following:

- oxygen saturation as a function of water temperature, salinity, and phosphate concentration; and
- oxygen saturation as a function of water temperature, salinity, phosphate concentration, and depth.

```{r}
# model 1
mdl1 <- lm(o2sat ~ t_deg_c + salinity + po4u_m, data = seawater)

# model 1 formula
f1 <- o2sat ~ t_deg_c + salinity + po4u_m

# model 2
mdl2 <- lm(o2sat ~ t_deg_c + salinity + po4u_m + depth_m, data = seawater)

# model 2 formula
f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m
  
```

### Comparing models with AIC

```{r}
# Calculating AIC
AIC(mdl1, mdl2)

# Table of corrected AIC values
aictab(list(mdl1, mdl2))
```

Based on these observation, Model 2 has better model fit (AIC = `r round(AICc(mdl2),2)`) compared to that of Model 1 (AIC = `r round(AICc(mdl1),2)`).

### K-fold cross validation of models

```{r}
folds <- 10
fold_vec <- rep(1:folds, length.out = nrow(seawater))
table(fold_vec)

set.seed(42)

seawater_fold <- seawater %>% 
  mutate(group = sample(fold_vec, size = n(), replace = FALSE))

# first fold
test_df <- seawater_fold %>% 
  filter(group == 1)
train_df <- seawater_fold %>% 
  filter(group !=1)
```

#### Root-mean-square error

```{r}
# go in reverse order of these operations. Find error, square it, find average, then take the square root.

calc_rmse <- function(x,y) {
  rmse_result <- (x-y)^2 %>% 
    mean() %>% 
    sqrt()
  return(rmse_result)
}
```


```{r}
# Using the training dataset above, create 2 linear regression models based on formula/function calc_rmse
training_mdl1 <- lm(f1, data = train_df)
training_mdl2 <- lm(f2, data = train_df)

#predict on test data

predict_test <- test_df %>% 
  mutate(model1 = predict(training_mdl1, test_df),
         model2 = predict(training_mdl2, test_df))

rmse_predict_test <- predict_test %>% 
  summarize(rmse_mdl1 = calc_rmse(model1, o2sat),
            rmse_mdl2 = calc_rmse(model2, o2sat))
```

```{r}
# calculating over all folds and taking the average

rmse_df <- data.frame()

for(i in 1:folds) {
  kfold_test_df <- seawater_fold %>% 
    filter(group == 1)
  kfold_train_df <- seawater_fold %>% 
    filter(group !=i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .))
  kfold_rmse <- kfold_pred_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, o2sat),
              rmse_mdl2 = calc_rmse(mdl2, o2sat))
  
  rmse_df <- bind_rows(rmse_df, kfold_rmse)
  
}

rmse_df_table <- rmse_df %>% summarize(mean_rmse_mdl1 = mean(rmse_mdl1),
                      mean_rmse_mdl2 = mean(rmse_mdl2))
```

### Conclusions

The root-mean-square-error (rmse) of Model 1 (rmse = `r round(rmse_df_table$mean_rmse_mdl1, 2)`) is slightly smaller than that of Model 2 (rmse = `r round(rmse_df_table$mean_rmse_mdl2, 2)`), making Model 1 the model of best fit following this assessment. (Data: CalCOFI 2021)

```{r}
final_mdl <- lm(f1, data = seawater)
summary(final_mdl)
```

Our final model:
`r equatiomatic::extract_eq(final_mdl, wrap = TRUE)`

`r equatiomatic::extract_eq(final_mdl, wrap = TRUE, use_coefs = TRUE)`


### Citations

CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.