---
title: 'Task 2: Model selection'
author: 'By: Claire Gonzales'
date: "1/24/2022"
output: 
  html_document: 
    theme: lumen
---

### Project Overview 


```{r setup, include= TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

# attach packages

library(tidyverse)
library(here)
library(lubridate)
library(AICcmodavg)
library(equatiomatic)
```

```{r}
# reading in csv data

seawater <- read_csv(here("data", "calcofi_seawater_samples.csv")) %>% 
  drop_na()
```

### Create 2 models

The two multiple linear regressiong models will address the following:

- oxygen saturation as a function of water temperature, salinity, and phosphate concentration; and
- oxygen saturation as a function of water temperature, salinity, phosphate concentration, and depth.

```{r}
# model 1
mdl1 <- lm(o2sat ~ t_deg_c + salinity + po4u_m, data = seawater)

# model 1 formula
f1 <- o2sat ~ t_deg_c + salinity + po4u_m

# model 2
mdl2 <- lm(o2sat ~ t_deg_c + salinity + po4u_m + depth_m, data = seawater)

# model 2 formula
f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m
  
```

### Comparing models with AIC

```{r}
# Calculating AIC
AIC(mdl1, mdl2)

# Table of corrected AIC values
aictab(list(mdl1, mdl2))
```

Based on these observation, Model 2 has better model fit (AIC = `r round(AICc(mdl2),2)`) compared to that of Model 1 (AIC = `r round(AICc(mdl1),2)`).

### Citations

CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.